Article figures

### 2020-12-01 src/article_figures.Rmd
Use the raw result dataframe as input for performance forest plot.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
save_figs = FALSE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

@@

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
# Give a more descriptive name to this variable
variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
```

```{r}
if (save_figs) {
    filename <- paste(fig_path, "effect_size_importance.pdf", sep="/")
} else {
    filename <- NULL
}
create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
```

@@

  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)

  var_imp <- varimp %>%
    mutate(Pretty=fct_rev(Pretty),
           mean=Importance,
           low=ifelse(is.na(Importance), NA, 0),
           high=Importance)

  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
    select(-Importance, -Variable, stripe)

@@

```{r}
res <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
g <- res$plot
final <- res$final
if (save_figs)
  ggsave(filename="alternative_effect_size_importance.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```

@@

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]
  s <- map(s, function(x) when(x, 
                               x == "progesa" ~ "eProgesa",
                               x == "finngen" ~ "Biobank",
                               x == "findonor" ~ "FinDonor",
                               ~ "koe"))
  return(s)
}

# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
  df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
  df <- bind_rows(df1, df2)
  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
  df <- df %>%
    mutate(Id=factor(Id, levels=rev(all_ids)),
           type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
           model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
    mutate(type = fct_recode(type, !!!new_names))

  thresholds <- get_threshold_df()
  df <- bind_rows(df, thresholds)
  df <- df %>% 
    mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
  write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")

  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(type %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, type=mytype) # Dummy table to use the same x-axis limits for E6 and E12

  dummy <- dummy %>%
    mutate(type = fct_recode(type, !!!new_names))
  print(dummy)
  print(df)

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Dummy"),
                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("type", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
  return(list(g=g, df=df))
}

@@

```{r}
res <- create_performance_forest_plot()
g <- res$g
df <- res$df
if (save_figs)
  ggsave(filename="performance_forest_plot.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```

@@

```{r}
g1 + theme_gray(base_size=20)
g2 + theme_gray(base_size=20)
```

```{r, eval=FALSE, include=FALSE}
jarkko_test <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")

  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)

  var_imp <- varimp %>%
    mutate(Pretty=fct_rev(Pretty),
           mean=Importance,
           low=ifelse(is.na(Importance), NA, 0),
           high=Importance)

  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
    select(-Importance, -Variable, stripe)

#   final <- tibble(
#             method = c("Regression coefficient",
#                        "Regression coefficient","Importance"),
#             Pretty = factor(c("Days to previous full blood donation","Days to previous full blood donation",
#                        "Days to previous full blood donation")),
#             #gender = c("male", "female", "both"),
#               mean = c(0.0152091681221149,
#                        0.0445741103338152,0.0255604042298423),
#                low = c(0.00691102953212081, 0.0353192308283623, 0),
#               high = c(0.0240855557569598,
#                        0.0537929726515129,0.0255604042298423)
# )

  #combined_variables <- tibble(Pretty=factor("Days to previous full blood donation"), stripe=factor(1))

  print(final)         
  print(combined_variables)
  vars <- c("Days to previous full blood donation", "Age", "Previous donation deferred", "Year")
  plot <- final %>% #filter(Pretty %in% vars) %>%
    ggplot(aes(color=gender)) +
    geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(#ymax = as.numeric(Pretty) + 0.5,
                          #ymin = as.numeric(Pretty) - 0.5
                          ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              fill = "blue", 
              #position="identity",
              xmin=-Inf, xmax=Inf, alpha = 0.1, inherit.aes = FALSE) +
    ggstance::geom_pointrangeh(#data=final, 
                               mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high),
                               #position=position_dodge2(0.8)
                               position=position_dodge2(width=0.8, padding=0.1)
                               ) +
    scale_y_discrete() +
    facet_wrap("method", scales = "free_x") +
    #coord_flip() +
    theme_classic(base_size = base_size)

  return(list(plot=plot, final=final))
}

res <- jarkko_test(male_posterior, female_posterior, variables, filename=filename)
final <- res$final
g <- res$plot
g
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
save_figs = TRUE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

@@

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
# Give a more descriptive name to this variable
#variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
variables <- recode(variables,    # Note that recode uses (stupidly) oldname = newname order, unlike many other functions in tidyverse
                    `Previous Hb deferral` = "Previous donation deferred",
                    `Polygenic risk score` = "Polygenic score",
                    `RNF43 mutant` = "RNF43")
```

```{r, eval=FALSE}
if (save_figs) {
    filename <- paste(fig_path, "old-effect-size-importance.pdf", sep="/")
} else {
    filename <- NULL
}
create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
```

@@

  # Get effect sizes of the linear model  
  dlmm <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
  dlmm <- dlmm %>%
    mutate(Pretty=fct_rev(Pretty))

  # Get variable importance from the RF object
  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
  var_imp <- varimp %>%
    mutate(Pretty=fct_rev(Pretty),
           mean=Importance,
           low=ifelse(is.na(Importance), NA, 0),
           high=Importance)

  final <- bind_rows(`Regression coefficient`=dlmm, Importance=var_imp, .id="method") %>%
    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
    select(-Importance, -Variable, stripe)

@@

```{r}
res <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
g <- res$plot
final <- res$final
if (save_figs)
  ggsave(filename="effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180)
  ggsave(filename="presentation-effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180, height=90)
g
```

@@

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
  # s <- map(s, function(x) when(x, 
  #                              x == "progesa" ~ "eProgesa",
  #                              x == "finngen" ~ "Biobank",
  #                              x == "findonor" ~ "FinDonor",
  #                              ~ "koe"))
  return(s)
}

# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  new_names <- c(`€/donation, 6 month deferral` = "E6", `€/donation, 12 month deferral` = "E12")
  if (FALSE) {
    df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
    df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
    df <- bind_rows(df1, df2)
    #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")

    df <- df %>%
      mutate(Id=factor(Id, levels=rev(all_ids)),
             type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
             model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
      mutate(type = fct_recode(type, !!!new_names))

    thresholds <- get_threshold_df()
    df <- bind_rows(df, thresholds)
    df <- df %>% 
      mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
    #write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
  } else {
#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")

    df <- df %>% 
      pivot_wider(names_from=c("type"), values_from=value) %>%
      mutate(low = ifelse(is.na(low), value, low),     # To make the plot work, copy low=high=value for thresholds
             high = ifelse(is.na(high), value, high))

    df <- df %>%
      mutate(Id=factor(Id, levels=rev(all_ids)),
             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
             model=case_when(variable == "threshold6" ~ "t6",
                             variable == "threshold12" ~ "t12",
                             TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")),
             variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
             variable = factor(variable, levels=c("AUROC", "AUPR", "F1", "E6", "E12", "Threshold")),
             variable = fct_recode(variable, !!!new_names))

  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(variable %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, variable=mytype) # Dummy table to use the same x-axis limits for E6 and E12

  dummy <- dummy %>%
    mutate(variable = fct_recode(variable, !!!new_names))
  print(dummy)
  print(df)

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Dummy"),
                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("variable", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  return(list(g=g, df=df))
}

@@

```{r}
res <- create_performance_forest_plot()
g <- res$g
df <- res$df
if (save_figs)
  ggsave(filename="performance-forest-plot.pdf", #  title="Performance forest plot", 
         path=fig_path, plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
g
```

@@

```{r}
base_size <- 20
presentation1 <- g1 + coord_fixed() + theme_gray(base_size=base_size)
presentation2 <- g2 + coord_fixed() + theme_gray(base_size=base_size)
if (save_figs) {
  ggsave(filename="presentation-classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=presentation1, dpi=600, units="mm", width=180, height=120)
  ggsave(filename="presentation-classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=presentation2, dpi=600, units="mm", width=180, height=120)
}

Pretty print results as a table.

```{r}
df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
df <- df %>% 
  pivot_wider(names_from=c("variable", "type"), values_from=value, names_glue="{variable} {type}")
names(df) <- str_remove(names(df), " value$")
df %>%
  select("Id", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
```

```{r, eval=FALSE, include=FALSE}
# jarkko_test <- function(male_posterior, female_posterior, variables,
#                                            filename=NULL,
#                                            width = 180,  # width of the combined figure in millimetres
#                                           base_size = 11
# ) {
#   rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
# 
#   combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
#     mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
#            Pretty=factor(Pretty, levels=Pretty))
# 
#   result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
#   varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
# 
#   result <- result %>%
#     mutate(Pretty=fct_rev(Pretty))
# 
#   var_imp <- varimp %>%
#     mutate(Pretty=fct_rev(Pretty),
#            mean=Importance,
#            low=ifelse(is.na(Importance), NA, 0),
#            high=Importance)
# 
#   final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
#     mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
#     select(-Importance, -Variable, stripe)
# 
# #   final <- tibble(
# #             method = c("Regression coefficient",
# #                        "Regression coefficient","Importance"),
# #             Pretty = factor(c("Days to previous full blood donation","Days to previous full blood donation",
# #                        "Days to previous full blood donation")),
# #             #gender = c("male", "female", "both"),
# #               mean = c(0.0152091681221149,
# #                        0.0445741103338152,0.0255604042298423),
# #                low = c(0.00691102953212081, 0.0353192308283623, 0),
# #               high = c(0.0240855557569598,
# #                        0.0537929726515129,0.0255604042298423)
# # )
#   
#   #combined_variables <- tibble(Pretty=factor("Days to previous full blood donation"), stripe=factor(1))
#   
#   print(final)         
#   print(combined_variables)
#   vars <- c("Days to previous full blood donation", "Age", "Previous donation deferred", "Year")
#   plot <- final %>% #filter(Pretty %in% vars) %>%
#     ggplot(aes(color=gender)) +
#     geom_rect(data=combined_variables %>% filter(stripe==1),
#               mapping=aes(#ymax = as.numeric(Pretty) + 0.5,
#                           #ymin = as.numeric(Pretty) - 0.5
#                           ymax = as.numeric(Pretty) + 0.5,
#                           ymin = as.numeric(Pretty) - 0.5),
#               fill = "blue", 
#               #position="identity",
#               xmin=-Inf, xmax=Inf, alpha = 0.1, inherit.aes = FALSE) +
#     ggstance::geom_pointrangeh(#data=final, 
#                                mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high),
#                                #position=position_dodge2(0.8)
#                                position=position_dodge2(width=0.8, padding=0.1)
#                                ) +
#     scale_y_discrete() +
#     facet_wrap("method", scales = "free_x") +
#     #coord_flip() +
#     theme_classic(base_size = base_size) 
# 
#   return(list(plot=plot, final=final))
# }
# 
# res <- jarkko_test(male_posterior, female_posterior, variables, filename=filename)
# final <- res$final
# g <- res$plot
# g
```
